Corporate Security Solutions for BYOD:\\ A Novel User-Centric and Self-Adaptive System

### 2015-07-09 Response to reviewers - r2.tex

Following the Editor's and some of the reviewers' requests, we have updated the state of the art, revising and commenting in the text (Section 2) several works with common topics with this manuscript. The new references can be consulted b the end of this letter.

% -----------------------------------------------------------------------------

\section{Comments by Reviewer \#2}

% Antonio - Voy a sustituir el Verbatim por un cambio de color en las
% respuestas porque al compilarlo no hace los saltos de lÌnea si no se
% los especificas y el texto se sale del margen. 
% Es quotation, no verbatim, øno?

@@

% -----------------------------------------------------------------------------

\section{Comments by Reviewer \#3}

\begin{verbatim}
This paper describes MUSES, an adaptable framework for enforcing 
security policies on mobile devices. This tool relies on Machine 
Learning and CI techniques to predict policy violations and is 
self-adaptive on user behaviors.
The idea seems interesting although several technical concerns 
arise from the reading of this work:
1.      The tools aims at enforcing security policies, however 
the security response occurs only AFTER a particular pattern is 
recognized to be potentially malicious. From my understandings, 
there is no a-priori enforcement except for the prediction based 
on event logs. This could mean that a security violation could 
in principle happen and only later recognized. This is not so 
suitable for a corporate environment with strict security 
requirements. This statement should be discussed and preferably 
discredited with experimental results.
\end{verbatim}

MUSES starts from a defined set of security rules that the CSO must
define. Ideally, these rules should deal, in principle, with all the
potential security incidents that the company aims to detect. MUSES
would be able to define new rules by means of refinement which could
be labelled, i.e. assigned a decision, after the corresponding events
have happened. However, it can also infer or create new rules, which
the CSO must approve, using computational intelligence
techniques. These rules could deal with unexpected situations, which
have not happened previously .

@@

Anyway, the idea is to define a set of `restrictive' rules at the very beginning, able to ensure the system safety in a big amount of situations. Then, MUSES will refine or improve this set, making more specific and optimal rules.

Moreover, MUSES offers visual analysis tools to the CSOs, which can
aid or guide them to detect suspicious or anomalous patterns, and to
study how to avoid these patterns becoming malicious to the company.

### 2015-07-09 MUSES_vs_BYOD_solutions.tex
Revised intro refs #7

The way how users employ their devices has changed in the past few
years; even the number of different devices per user is increasing
because of the so-called Internet of Things
\cite{weber2010internet}. Thanks to it, users have more flexibility,
and also at their works, because now they can use their devices to
access company assets and continue working outside of the
workplace. To allow this way of working, companies started to adopt
what has been called the Bring Your Own Device (BYOD) philosophy.

Data security and privacy are key factors for a company, and thus, in
order to protect them, it is usual that the Chief Security Officer (CSO) of a company defines a set of Organizational Security
Policies \cite{Opp_Security11}. However, allowing a BYOD environment means to take care of the risks that the use of personal devices can bring to the company \cite{gangula2013survey}. For instance, the mixture between personal and professional information in these devices would allow the user to navigate inside social networks, where important data could be leaked in the event of a security incident.

In this scenario several solutions have appeared in order to manage corporate security. Most of the solutions are focused for both smartphones and desktop PCs, although some are implemented for a certain platform. However, almost all of them try to be non-intrusive -with regard to users' personal data-, friendly, and easy to use.

Moreover, it has been demonstrated that employees who ignore company security policies are the main hazard for company security \cite{Adams_Users05}. In this situation, the tools that are appearing in the market aim to cope with the concept of seamless working experience on different devices. This concept is a methodology of work which allows users to start, or continue a working session, over multiple devices and locations, without any significant loss of data. This new situation has a big impact from the point of view of the security \cite{Schu_SecPatterns05}, since company data borders have changed in the last years. Therefore, now the users can access important data from outside the enterprise, and possibly through a non completely secure channel. %Antares - Yo quitaba lo de significant loss of data, ya que significant se repite después.
%Cambiado

@@

The general ideas behind MUSES were already presented in
\cite{DBLP:conf/sac/MoraCGZJEBAH14,DBLP:conf/gecco/MoraCGZE14}. However,
the implementation was in its early stages when those papers were
published. Also, an additional work on applying data mining and machine learning techniques was conducted in \cite{mora14:urls}. It presented a study over a dataset similar to those that MUSES manages, but not the same, since the prototype of MUSES had not still been released.
Thus, a set of URL connection logs were provided by a Spanish security company, % [Pedro] quizás decir que es una empresa de seguridad o algo así % Añadido, gracias!
as well as both black and white lists. These lists contained the URLs that were forbidden, or allowed, to be accessed by the employees. 
By taking the lists as `rules', and the HTTP connections as `events', a previous simulation of the rule generation through classification, and its results, were presented in that work. % [Pedro] ¿ese "could be" se refiere a verbo en pasado? en ese caso, quizás usar "... was presented at..." % Gracias again

@@

The rest of the paper is organized as follows. First, some background situation about enterprise security is presented in Section \ref{sec:preliminaryconcepts}, explaining how BYOD philosophy affects it. The main solutions or tools are detailed in Section \ref{sec:toolsreview}, even if they are still in development.  
Section \ref{sec:muses} presents the Multi-platform Usable Endpoint Security System, with its architecture and the soft computing techniques it makes uses of. Advantages and benefits of the proposed system are compared to other solutions found in bibliography in Section \ref{sec:comparison}.
Finally, the reached conclusions are discussed in Section \ref{sec:conclusions}, followed by future work.

>>>>

The way how users employ their devices has changed in the past few
years; even the number of different devices per user is increasing
because of the so-called Internet of Things
\cite{weber2010internet}. Thanks to it, users have more flexibility
since they can now use their devices to
access company assets and continue working outside of the
workplace. To allow this way of working, companies have started to adopt
what has been called the Bring Your Own Device (BYOD) philosophy.

But data security and privacy are key factors for a company, and thus, in
order to protect them, it is usual that the Chief Security Officer (CSO) of a company defines a set of Organizational Security
Policies \cite{Opp_Security11}; allowing a BYOD environment implies
taking care of the risks that the use of personal devices can bring to
the company \cite{gangula2013survey}. For instance, the mixture
between personal and professional information in these devices would
allow the user to navigate inside social networks, where important
data could be leaked in the event of a security incident.

In this scenario several solutions that manage corporate security have
appeared. Most of them are geared for both smartphones and desktop
PCs, although some are implemented only for a certain platform. However,
almost all of them try to be non-intrusive with regards to users'
personal data, friendly, and easy to use.

Even with the help of those solutions, it has been demonstrated that
employees who ignore company security policies are the main hazard for
company security \cite{Adams_Users05}. In this situation, the tools
that are appearing in the market aim to cope with the concept of
seamless working experience on different devices. This concept is a
methodology of work which allows users to start, or continue a working
session, over multiple devices and locations, without any significant
loss of data. This new situation has a big impact from the point of
view of the security \cite{Schu_SecPatterns05}, since company data
borders have changed in the last years. Therefore, now the users can
access important data from outside the enterprise, and possibly
through a non completely secure channel. %Antares - Yo quitaba lo de
                                %significant loss of data, ya que
                                %significant se repite después. 
%Cambiado

@@

The general ideas behind MUSES were already presented in
\cite{DBLP:conf/sac/MoraCGZJEBAH14,DBLP:conf/gecco/MoraCGZE14}. However,
the implementation was in its early stages when those papers were
published. Also, an additional work on applying data mining and
machine learning techniques was conducted in \cite{mora14:urls}. This
last paper presented a study over a dataset similar to those that MUSES manages, but not the same, since the prototype of MUSES had not still been released.
Thus, a set of URL connection logs were provided by a Spanish security company, % [Pedro] quizás decir que es una empresa de seguridad o algo así % Añadido, gracias!
as well as both black and white lists. These lists contained the URLs that were forbidden, or allowed, to be accessed by the employees. 
By taking the lists as `rules', and the HTTP connections as `events', a previous simulation of the rule generation through classification, and its results, were presented in that work. % [Pedro] ¿ese "could be" se refiere a verbo en pasado? en ese caso, quizás usar "... was presented at..." % Gracias again

@@

The rest of the paper is organized as follows. First, some background
material about enterprise security is presented in Section
\ref{sec:preliminaryconcepts}, explaining how the BYOD philosophy
affects it. The main solutions or tools available or in development
are detailed in Section \ref{sec:toolsreview}.
Section \ref{sec:muses} presents the Multi-platform Usable Endpoint
Security System, with its architecture and the soft computing
techniques it makes uses of. Advantages and benefits of the proposed
system are compared to other solutions found in bibliography in
Section \ref{sec:comparison}. 
Finally, the reached conclusions are discussed in Section
\ref{sec:conclusions}, followed by future work.
